\graphicspath{{Figures/Chameleon}}

\chapter{\chameleon{}: Interactive Type Error Exploration}


\label{chap:chameleon} 

This chapter explores the challenges associated with traditional tools for debugging type errors in Haskell. We focus on their limitations in identifying all relevant locations, explaining errors in user-friendly language, and the tendency to introduce bias. We introduce \chameleon{}, a type error debugging tool aimed at exploring multi-step type errors efficiently and intuitively. \chameleon{} addresses the common challenges of traditional compiler tools and presents type errors in an optimized, comprehensible manner. We delve into the design of \chameleon{} and demonstrate its usage through examples of resolving type errors. Finally, we present our user studies designed to evaluate the effectiveness of type error visualization and interactive debugging tools.



 This chapter uses the content from our paper \textit{ChameleonIDE: Untangling Type Errors Through Interactive Visualization and
Exploration} \cite{Fu2023-fm}, with slight modifications to ensure coherence within this thesis. This work was published and presented at the International Conference on Program Comprehension (ICPC) 2023.

\input{Chapters/chameleon-sections/introduction.tex}
\input{Chapters/chameleon-sections/chameleon.tex}
\input{Chapters/chameleon-sections/walkthrough.tex}
\input{Chapters/chameleon-sections/evaluation.tex}
\input{Chapters/chameleon-sections/discussion.tex}
\input{Chapters/chameleon-sections/related-work.tex}
\input{Chapters/chameleon-sections/conclusion.tex}